---
title: "Data Wrangling 1"
author: "Georgie Knight"
date: "8 August 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The task here is to load a data set which is saved as a .csv file and 'clean up' the data in order to make it easier to analyse.

## Dplyr

We first load the _dplyr_ and _tidyr_ packages which will help us wrangle the data:

```{r}
library("dplyr")
library("tidyr")
```
## Loading up the data

Then we load the file 'refine_original.csv' into R and save it as a data frame called 'data_fr':

```{r}
refine_original <- read.csv("C:/Users/Georgie/Dropbox/Springboard/DataWrangling1/refine_original.csv")
data_fr <- data.frame(refine_original)
```

We then convert it to a table  called 'my_tbl' within the dplyr package...

```{r}
my_tbl<-dplyr::tbl_df(data_fr)

```
... and take a quick look at it:
```{r, echo = FALSE}
my_tbl
```

##Task 1: Clean up brand names

_Clean up the 'company' column, so all of the misspellings of the brand names are standardized._

We first standardise the column company by using the _mutate_ function combined with the _tolower_ function

```{r}
my_tbl <- mutate(my_tbl, company = tolower(company))
my_tbl$company 
```

Check the first letters of the company names that are unique as we will use these to rewrite the company column:

```{r}
temp <- my_tbl %>% select(company) %>% mutate(first_letter = substr(company,1,1))
unique(temp$first_letter)
```
We then use a _pipe_ to first create a column which contains the first letter of the company name and then use this column along with _replace_ to standardise all the names


```{r}
my_tbl <- my_tbl  %>% 
  mutate(first_letter = substr(company,1,1))  %>% 
  mutate(company = replace(company, first_letter == "p", "philips")) %>% 
  mutate(company = replace(company, first_letter == "a", "akzo")) %>% 
  mutate(company = replace(company, first_letter == 'f', 'philips')) %>% 
  mutate(company = replace(company, first_letter == "v", "van houten")) %>%
  mutate(company = replace(company, first_letter == 'u', 'unilever')) %>% 
  select(-first_letter)

my_tbl$company
```

##Task 2: Separate product code and number

_Separate the product code and product number into separate columns_

We'll use the _separate_ function from _tidyr_ to do this.
```{r}
my_tbl <- 
separate(my_tbl, Product.code...number , c("product_code", "product_number"),
         sep = '-', remove = FALSE)
my_tbl
```

##Task 3: Add product categories
_You learn that the product codes actually represent the following product categories: p = Smartphone, v = TV, x = Laptop, q = Tablet. In order to make the data more readable, add a column with the product category for each record._

We again just use the _mutate_ function along with _replace_ for this.
```{r}
my_tbl <- my_tbl  %>% 
mutate(product_category = product_code)  %>% 
mutate(
product_category = replace(product_category, product_code == "p", "Smartphone"), 
product_category = replace(product_category, product_code == "v", "TV"),
product_category = replace(product_category, product_code == "x", "Laptop"),
product_category = replace(product_category, product_code== "q", "Tablet")
      ) 

select(my_tbl,product_code, product_category)
```


##Task 4: Add address

Concatenate the address, city and country into a full_address column using the _paste_ function:

```{r}
my_tbl <-
  mutate(
         my_tbl, full_address = paste(address, city, country, sep= ',')
         )

my_tbl$full_address
```  

##Taks 5: Create dummy variables for company and product category
For each company and product, create a binary $0,1$ column which contains $1$ when the company/ product is in that row, $0$ otherwise. We use a TRUTH assignment converted to integer for this.

```{r}
my_tbl <-mutate( my_tbl,
          company_philips = as.integer(company == 'philips'), 
          company_akzo = as.integer(company == 'akzo'),
          company_van_houten = as.integer(company == 'van houten'),
          company_unilever = as.integer(company == 'unilever'),
          product_smartphone = as.integer(product_code == 'p'), 
          product_tv = as.integer(product_code == 'v'),
          product_laptop = as.integer(product_code == 'x'),
          product_tablet =  as.integer(product_code == 'q')
          )

select(my_tbl, contains('com'))
select(my_tbl, contains('product_'), -product_number)
```

Now save the cleaned table as a .csv.
```{r}
write.csv(my_tbl, file="refine_clean.csv")
```  
  